XOR-Some Prolific Pilot 2

### 2021-04-04 analysis/xor-some-prolific-pilot2.Rmd
add stats

cat("Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
cat("Gender counts: ") 
d_native %>% count(gender) %>% mutate(n = n/88)
```

@@

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_long %>% 
  filter(block == "xor" | block == "some") %>%
  #group_by(block, class_condition, prior_class) %>%
  #summarize(mean_rating = mean(response)) %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line() +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Mean inference strength ratings by-predictor")

Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.

@@

@@

``` {r}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

@@

## Stats

```{r, results='hide'}
# xor, maximal model with interactions and maximal REs
model_xor <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@

```{r, results='hide'}
# some, maximal model with interactions and maximal REs
model_some <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some)
```

## Explore z-scoring

## Exploreusing binary predictors

>>>>

cat("Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
d_native %>% count(gender) %>% mutate(n = n/88)
```

@@

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_long %>% 
  filter(block == "xor" | block == "some") %>%
  #group_by(block, class_condition, prior_class) %>%
  #summarize(mean_rating = mean(response)) %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

```
Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.

@@

```{r}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```

Plot inference strength ratings against raw predictor ratings for each participant / each item:
```{r}
d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

d_critical_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp

d_critical_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```

@@

``` {r}

@@

## Explore z-scoring
Visual exploration of z-scoring each measure (rel / comp / pri / target / comp-wUtterance / rel-wUtterance) per participant, collapsing some and or together. 
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd)

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response_centered)) -> d_critical_zScore_summary

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
  ggtitle("Centered responses to by-predictor questions")
```

Plot z-scored predictor ratings against z-scored target ratings
```{r}
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

d_critical_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri.z

gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
```

## Stats
Maximal models on raw ratings:
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs
model_xor <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@

```{r, results='hide'}
# some, maximal model with interactions and maximal REs
model_some <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some)
```

#### Stats on z-scored data
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs on z-scored data
model_xor_zScored <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored)
```

```{r, results='hide'}
# some, maximal model with interactions and maximal REs
model_some_zScored <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored)
```

#### Stats with dichotomized predictors 
Maximal model on raw inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = as.factor(relevance),
         competence = as.factor(competence),
         prior = as.factor(prior)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat
# sum code predictors
# low prior: 1, high prior:: -1 
contrasts(d_critical_wide_cat$prior) <- contr.sum(2)
# low comp : 1, high comp : -1
contrasts(d_critical_wide_cat$competence) <- contr.sum(2)
# low rel: 1, high rel: -1
contrasts(d_critical_wide_cat$relevance) <- contr.sum(2)

# xor, maximal model with interactions and maximal REs
model_xor_cat <- brm(
  response ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_cat)
```

```{r, results='hide'}
model_some_cat <- brm(
  response ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat)
```

Maximal model on z-scored inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = as.factor(relevance),
         competence = as.factor(competence),
         prior = as.factor(prior)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat_zScored
# sum code predictors
# low prior: 1, high prior:: -1 
contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
# low comp : 1, high comp : -1
contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
# low rel: 1, high rel: -1
contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)

# xor, maximal model with interactions and maximal REs
model_xor_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_cat_zScored)
```

```{r, results='hide'}
model_some_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat_zScored)
```

In summary, the different models provide somewhat different results, specially when using raw ratings vs. binary predictors. Since interpreting the categorical models in relation to the plots might be not as straight-forward, I am not entirely sure whether the qualitative results match in all models. Using centered data seems to provide a little tighter estimates in the models; not sure where the slight qualitative differences come from.
